---
title: "tidychess"
author: "Josh Merrell"
output: 
  github_document:
    toc: true
---
# Introduction

**Terminology**  

  * __game data__  
  * __positional data__  
  * __PGN format__  
  * __FEN format__  
  * __matrix format__  
  * __tidy format__  
  * __positional mobility__  
  * __pathPrior__: planning to change the name to _theoretical mobility_. Lists the squares a piece can move on an empty chessboard.  
  * __Post functions__: planning to change the name to _ingame_ functions. They were named for _aposteriori_, and they list the squares to which a piece can move from a given position.  


# Exploration of Functions
**Setup**
```{r setup}
invisible(lapply(list.files("R", "\\.R$", full.names = T), source))
```

**The mobility_piece function**
```{r mobility_piece}
# if position_vec is not passed as an input, 
# it is assumed that no other pieces are on the board
mobility_piece(piece = "bishop", square = "e5")

# if board is empty, specified piece is not on specified square:
mobility_piece(piece = "white pawn", square = "a2", position_vec = empty())

# white pawns travel up and black pawns travel down
mobility_piece(piece = "black pawn", square = "e5")
mobility_piece(piece = "white pawn", square = "e5")

# pawns of unspecified color might go either direction
mobility_piece(piece = "pawn", square = "e5")

# mobility_piece() piece type from position_vec & square:
mobility_piece(square = "a2", position_vec = setup(includeEmpty = F)[1,])

# if piece input conflicts with square & position_vec inputs, it is overriden:
mobility_piece(square = "a2", 
               position_vec = setup(includeEmpty = F)[1,], 
               piece = "white rook")
```
**compare post & prior**
```{r mobility_board, eval=T}
mobility_board(position_vec = setup())
patterns <- c("white bishops" = "white bishop",
              "black knights"="black knight",
              "pawns"="pawn")
pieces <- unique(unlist(setup()[2,]));  pieces <- pieces[!is.na(pieces)]
mobility_board(position_vec = setup(), patterns = patterns)
mobility_board(position_vec = setup(), patterns = pieces)
rm(patterns, pieces)
```

**new_position**
```{r new_position, eval=F}
positions <- setup()
positions <- rbind(positions, new_position(new_pgn = "1.e3",
                                           position_vec = positions))
positions[,c("e2","e3")]
rm(positions)
```

**to.FEN**
```{r to.FEN, eval=F}
to.FEN(setup()[1,])
```

# Chess Data
**Bash-managed scraping**
```{r bash-scrape, engine="bash", eval=F}
#!/bin/bash
# obtain list of previously stored data to avoid duplication
aws s3 ls s3://jbchess/data --recursive > temp.txt

# iterate through a range of numbers
for n in $(seq 200 1000)
do

  # adjust both numbers to the proper order of magnitude
	first=$(($n*1000+1))
	last=$(($n*1000+1000))
	# record the beginning & ending range of numbers; include a timestamp
	echo start $first $last $(date)
	
	# pass the range of numbers to the scrape.chess.com function
	Rscript R/scrape.chess.com.R $first $last
	
	# move the acquired data to the bucket
	aws s3 mv data s3://jbchess/data --recursive
done
```

**acquire sample of chess data**
```{r s3-retrieve, engine="bash", eval=F}
# use the shell to extract sample of chess games:
aws s3 cp "s3://jbchess/data/chess.com IDs 1-1000.csv" "data/chess.com IDs 1-1000.csv"
```

**build data frame of chess positions**
```{r positional data}
DATA <- read.csv("data/chess.com IDs 1-1000.csv", stringsAsFactors = F)
DATA <- DATA[grepl(".",DATA$pgn),]
DATA[1,]
positions <- setup()
pgn <- strsplit(DATA[1,"pgn"]," ")[[1]]
pgn <- pgn[-length(pgn)]
pgn[seq(2,length(pgn),2)] <- paste0(seq(1,length(pgn)/2),
                                    "...",
                                    pgn[seq(2,length(pgn),2)])
for(elem_p in pgn){
  positions <- rbind(positions,
                     new_position(elem_p, positions))
}
apply(positions[-1,], 1, mobility_board, 
      patterns = c(white="white", black="black", pawns="pawn", 
                   knights="knight", bishops="bishop", rooks="rook", 
                   queens="queen", kings="king"))
```
ISSUE: phantom black pawn not being deleted
